workout_probabilities

### 2019-04-29 workout_probabilities.Rmd
starting to work on fdr

>>>>

# False positive rate control.

We develop a method, following Efron(2008), for controlling the false positive rate.
Every site with sufficient coverage and at least 1 alternate read falls into one of two classes, they are either *null* (non-variant with $\nu = 0$) or *nonnull* (variant with $\nu = \hat{f}$) with prior probabilities $p_0$ and $p_1 = 1-p_0$,

$$
\begin{array}{ll}
p_0 = \textrm{P}\{\textrm{null}\} \quad & \textrm{f}_{0}(\mathbf{x}) \quad \textrm{density if null}\\
p_1 = \textrm{P}\{\textrm{nonnull}\} \quad & \textrm{f}_{1}(\mathbf{x}) \quad \textrm{density if nonnull} .
\end{array}
$$

In this application we can a compute an estimate of $p_0$ as the ratio of the expected number of false positive sites and the number of expected true positive sites with variant allele frequency $\nu = \hat{f}$.
At a given site, with read depth $D$, the probability of seeing $n$ reads in error is binomial
$$
p_e \sim \textrm{Bin}(n,D,p=\bar{q})
$$
where $\bar{q}$ is the average quality score of the alternate reads.
(This is an estimate, they are have different q-scores)
The expected number of false positives with probability $p_e$ is then $N*p_e$, where $N$ is the number of sites sequenced.
In order to compute the expected nubmer of true positives with a given vaf, we need to make an assumption about the allele frequency distribution. 
In a neutrally evolving tumor, the count of variants with a given allele frequency is(bozic et al)
$$
N(f) = \frac{N\mu}{f}.
$$
